\begin{abstract}
\noindent
Does it matter how households value publicly provided goods relative to private
consumption, and does this depend on where they sit in the income distribution?
We show that it does -- and that both questions have first-order implications
for the size of the fiscal multiplier. Embedding \EC{} between public and private
consumption into a tractable heterogeneous-agent New Keynesian (THANK) model, we
derive a closed-form fiscal multiplier and characterize two new channels absent
from standard frameworks. The first is a \emph{direct fiscal demand} channel
through which government spending stimulates private consumption above and
beyond standard income effects. The second emerges when poorer constrained
households treat public goods as complements while wealthier households
substitute privately -- a configuration strongly supported by evidence that
public education, healthcare, and transport disproportionately benefit
low-income households. This distributional asymmetry generates curvature
heterogeneity across household types that \emph{decouples income cyclicality
from the hand-to-mouth share}, providing a new route to the MPC/MPE trilemma
that operates through the government-spending margin rather than through labor
supply frictions. A distributional dominance condition governs the aggregate
multiplier: fiscal policy is most effective precisely in the economies where it
is most needed -- those with high inequality and strong complementarity between
public services and the private consumption of the poor. Taking the model to a
quantitative HANK calibration disciplined by US micro data, we show that
plausibly heterogeneous internalization raises impact multipliers by
0.12 to 0.34 log points relative to the separable benchmark, with
welfare gains from targeted fiscal expansion of up to 1.8\% of steady-state
consumption for the bottom quintile. The framework provides a transparent
mapping from the distributional incidence of public goods to aggregate fiscal
transmission, with direct implications for developing-economy policy design.
\end{abstract}
